Plant species traits and growth rates: meta-analysis shows correlations change with plant size as predicted

### 2015-08-07 MS.tex
Add Script for publication biais

\item
Growth rates have often been seen as central elements of plant ecological strategies, and as linked to species traits. Yet, the literature is inconsistent about empirical correlations between functional traits and plant growth rate (GR), casting doubt on the capacity of traits to predict growth.

@@

\section*{Introduction}\label{introduction}
The phrase "plant ecological strategy" refers to how species face the challenge they all share:  to acquire sufficient water, nutrients and light for individual growth and to live long enough to replace themselves with offspring. An idea that species strategy differences might involve fast vs slow growth goes back at least to Grime and Hunt \citeyearpar{Grime:1975gr}, who measured a wide spread of relative growth rates for seedlings during early exponential growth under favourable conditions. Seedling potential relative growth rate (RGR, the change in a mass/height or diamater per time interval relative to initial size) has continued to be used as a major strategy indicator within the CSR (Competitor - Stress tolerator - Ruderal) scheme (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into its components (see equation X in Appendix 1) showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx}. From the 1990s a new style of "trait ecology" arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size Inquist et al () \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see Text Box 1\ref{sec:growth}).

@@

The main effect of increased seed size is to produce a larger overall seedling. Generally relative growth rate decreases with size (REF Enquist et al 1999 Nature, Tilman 1988?), and as such larger seedlings are predicted to have a lower relative growth rate during the earliest phases of growth \citet{Turnbull:2012ew}. However, among larger plants we expect the effect of this initial size to be minimal, and thus the correlation between seed size and growth rate should disappear.

@@

\textbf{Effect on absolute (AbsGR) and relative growth rate (RGR}
Tree growth is the increase in dimensions (i.e. diameter, height or mass) of an individual tree through time. While plant functional traits measurements are standardized across plant ontogeny, plant growth is commonly expressed differently according to the plant size/stage considered. Thus, mass and height are both used for smaller plants, whereas height and diamater are more convenient measures for larger tree. In addition, the growth increment (i.e. rate of change of a dimension per unit time) can be expressed in either absolute or relative terms. Relative growth rate (RGR) is commonly used for smaller plants, but is less useful for larger-plants because there is a very rapid decline in RGR with increasing tree size due to accumulation of support tissues (bark, branches etc). According to the growth model used here, the correlation between trait and growth is predicted to revert according to the growth measurement used for both SLA and seed mass. The predicted relationship between growth and either SLA or seed mass in seedlings differs if absolute growth rate in mass is used instead of $RGR_{mass}$ (i.e. becomes negative for SLA, positive for SM).

@@

Following recent call for the adoption of a checklist of quality criteria for meta-analysis in ecology (Koricheva and Gurevitch, 2014), we conducted several analyses to assess the reliability of our meta-analysis. Table A2 in Appendix X reported how and where (i.e in the text and/or figures) we addressed the 14 methodological quality criteria listed by Koricheva and Gurevitch (2014). In particular, we checked the dataset for publication bias with funnel plots (\citealt{Koricheva:2013tz}) and the 'trim and fill' method (Appendix), we explored the temporal changes in effect size (Appendix), and we estimated the extend of heterogeneity using the statistic I2 (Appendix X).

@@

Finally, we met all the quality criteria required for meta-analysis in ecology. We found no evidence of publication bias, the extend of the heterogenity appares to be low to moderate in our meta-analysis, and there is no temporal change in effect size calculated within a stage.

@@

\begin{table}[ht]
\centering
\caption{\textbf{Checklist of quality criteria for meta-analysis in ecology developed by Koricheva and Gurevitch (2014)} 
\label{TableA2}
\vspace{0.5cm}
\begin{tabular}{p{3cm}p{3cm}p{8cm}}
  \hline
Criteria & Addressed & \\
  \hline
Aarea & Photosynthetic rate per area & "photosynthetic rate per area" OR "rate of CO2 assimilation per area" OR "Aarea" OR "Amax" OR "photosynthetic capacity per area" \\
  Amass & Photosynthetic rate per mass & "photosynthetic rate per mass" OR "rate of CO2 assimilation per mass" OR "Amass" OR "Amax" \\
  Hmax & Maximum height & "potential height" OR "maximum height" OR "Hmax" \\
  Ks & Stem specific conductance & "Ks" OR "stem specific conductance" OR "stem hydraulic conductance" OR "sapwood conductance" \\
  LA & Leaf lamina area & "leaf length" OR "leaf lamina area" OR "leaf width" OR "leaf size" \\
  LAR & Leaf area ratio & "leaf area ratio" \\
  LMR & Leaf mass ratio & "leaf mass ratio" OR "leaf mass fraction" \\
  NARarea & Net assimilation rate per area &  "net assimilation rate per area" OR "rate of dry mass increament per area" OR "NARarea" OR "NAR" \\
  Narea & Leaf nitrogen content per area & "Narea" OR "nitrogen per area" OR "N/area" OR "leaf nitrogen content per area" OR "LNCarea" OR "leaf nitrogen content" \\
  NARmass & Net assimilation rate per mass &  "net assimilation rate per mass" OR "rate of dry mass increament per mass" OR "NARmass" OR "NAR" \\
  Nmass & Leaf nitrogen content per mass & "Nmass" OR "nitrogen per mass" OR "N/mass" OR "leaf nitrogen content per mass" OR "LNCmass" OR "leaf nitrogen content" \\
  Parea & Leaf phosphorus content per area & "Parea" OR "leaf P" OR  "leaf phosphorus content per area"  OR "leaf phosphorus content" \\
  Pmass & Leaf phosphorus content per mass & "Pmass" OR "leaf P" OR  "leaf phosphorus content per mass" OR "leaf phosphorus content" \\
  SA/LA & Sapwood area per leaf area & "sapwood area per leaf area" OR "huber value" OR "leaf area per sapwood area" \\
  Seed mass & Seed mass & "seed mass" OR "seed size" OR "seed volume" \\
  SLA & Specific Leaf Area & "leaf mass per area" OR "specific leaf area" OR "leaf construction cost" \\
  Thickness & Leaf thickness & "leaf thickness"  OR "leaf tissue density" \\
  Vessel density & Vessel density & "vessel density" OR "stem conduit density" \\
  Vessel size & Vessel area or diameter & "conduit area" OR "stem conduit area" OR "vessel diameter" OR "vessel size" OR "conduit size"   \\
  WD & Wood density & "wood density" OR"WD" OR "wood specific gravity" OR "stem density" OR "stem specific density" OR "SSD" \\
   \hline
\end{tabular}
\end{table}

% we conducted two types of analyses to assess the reliability of our meta-analysis. First, we estimated heterogeneity using a modified version of the heterogeneity statistic I2 (originally defined by Higgins & Thompson, 2002; see Supporting information for details). Heterogeneity estimates the consistency of the results obtained from published stud- ies used in our meta-analysis. A low level of heteroge- neity means that discrepancies among the effects reported by the original studies are satisfactorily explained solely by sampling errors and are not the result of different methodologies or species characteris- tics (Higgins & Thompson, 2002; Higgins et al., 2003; Cooper & Hedges, 2009). Second, we assessed our data set for signs of publication bias by visually inspecting funnel plots (Sterne et al., 2005) and also by conducting Egger’s regressions (Egger et al., 1997; Peters et al., 2006). We applied an Egger’s regression design, modi- fied for lnOR to data points consisting of the residuals of the effect sizes plus their measurement errors
% a 2012 THE AUTHORS. J. EVOL. BIOL. 25 (2012) 1911–1917
% JOURNAL OF EVOLUTIONARY BIOLOGY a 2012 EUROPEAN SOCIETY FOR EVOLUTIONARY BIOLOGY
% Meta-analysis of cost of parental care in birds 1913
% 1914 E. S. A. SANTOS AND S. NAKAGAWA
% (Peters et al., 2006; see Supporting information for details). If the intercept of the Egger’s regression is sig- nificantly different from zero, it is evidence of p

@@

\begin{figure}[htbp]
\centering
\includegraphics{output/FigA9.pdf}
\caption{\textbf{Funnel plots showing transformed correlation coefficient z between growth and the traits a) Specific leaf area (SLA), b) Wood density (WD), c) Asymptotic height (Hmax), d) Seed mass and e) Assimilation rate of CO2 (Aarea)}. Shown here are complete dataset (recorded at both interspecific and intraspecific level and under unstressed and stressed conditions). Point colors: red for adult, orange for sapling, grey for juvenile. Note that in these plots the x axis is the number of species included in the correlation, not the replication for the correlation. The number of species is sometimes as low as 1, and these are within-species correlations. The number of replicates used for the correlation coefficient was often much greater than the number of species. Nevertheless, the characteristic shape of funnel plots is observed, with correlations involving few species showing much wider variation even though within-species replication was sometimes high.}

@@

\begin{figure}[htbp]
\centering
\includegraphics{output/FigA10.pdf}
\caption{\textbf{Growth measurement used by traits and by size.} Shown here are complete dataset (recorded at both across and within species and under unstressed and stressed conditions). The colors correspond to the stage for which the r values have been recorded: red bars corresponds adult stage, orange bars to saplings, and grey bars to juveniles.}
\label{FigA10}
\end{figure}

>>>>

\item
Growth rates (GR) have often been seen as central elements of plant ecological strategies, and as linked to species traits. Yet, the literature is inconsistent about empirical correlations between functional traits and plant growth rate, casting doubt on the capacity of traits to predict growth.

@@

\section*{Introduction}\label{introduction}
The phrase "plant ecological strategy" refers to how species face the challenge they all share:  to acquire sufficient water, nutrients and light for individual growth and to live long enough to replace themselves with offspring. An idea that species strategy differences might involve fast vs slow growth goes back at least to Grime and Hunt \citeyearpar{Grime:1975gr}, who measured a wide spread of relative growth rates for seedlings during early exponential growth under favourable conditions. Seedling potential relative growth rate (RGR, the change in a mass/height or diamater per time interval relative to initial size) has continued to be used as a major strategy indicator within the CSR (Competitor - Stress tolerator - Ruderal) scheme (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into its components showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx}. From the 1990s a new style of "trait ecology" arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Enquist:2007ek,Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see Text Box 1\ref{sec:growth}). Today, we have no

@@

The main effect of increased seed size is to produce a larger overall seedling. Generally relative growth rate decreases with size, and as such larger seedlings are predicted to have a lower relative growth rate during the earliest phases of growth \citet{Turnbull:2012ew}. However, among larger plants we expect the effect of this initial size to be minimal, and thus the correlation between seed size and growth rate should disappear.

@@

\textbf{Effect on absolute (AbsGR) and relative growth rate (RGR}
Tree growth is the increase in dimensions (i.e. diameter, height or mass) of an individual tree through time. While plant functional traits measurements are standardized across plant ontogeny, plant growth is commonly expressed differently according to the plant size/stage considered. Thus, mass and height are both used for smaller plants, whereas height and diamater are more convenient measures for larger tree. In addition, the growth increment (i.e. rate of change of a dimension per unit time) can be expressed in either absolute or relative terms. Relative growth rate (RGR) is commonly used for smaller plants, but is less useful for larger-plants because there is a very rapid decline in RGR with increasing tree size due to accumulation of support tissues (bark, branches etc). According to the growth model used here, the correlation between trait and growth is predicted to revert according to the growth measurement used for both SLA and seed mass (see Text Box 1\ref{sec:growth}. The predicted relationship between growth and either SLA or seed mass in seedlings differs if absolute growth rate in mass is used instead of $RGR_{mass}$ (i.e. becomes negative for SLA, positive for SM).

@@

Following recent recommendations for the adoption of a checklist of quality criteria for meta-analysis in ecology (\citealt{Koricheva:2014ku}), we conducted several analyses to assess the reliability of our meta-analysis. Table A2 in Appendix 1 reported how and where (i.e in the text, appendix and/or figures) we addressed the 14 methodological quality criteria listed by \citet{Koricheva:2014ku}.

@@

Finally, we met all the quality criteria required for meta-analysis in ecology. We found no evidence of publication bias (). The extend of the heterogenity was high in our meta-analysis, and substantially explained by the plant stage for SLA and Seedmass as expected. Finally, we showed no temporal change in effect size for all the traits.

@@

\begin{table}[ht]
\centering
\caption{\textbf{Checklist of quality criteria for meta-analysis in ecology developed by Koricheva and Gurevitch (2014)} }
\label{TableA2}
\vspace{0.5cm}
\begin{tabular}{p{3cm}p{3cm}p{8cm}}
  \hline
Criteria & Addressed & \\
  \hline
Aarea & Photosynthetic rate per area & "photosynthetic rate per area" OR "rate of CO2 assimilation per area" OR "Aarea" OR "Amax" OR "photosynthetic capacity per area" \\
  Amass & Photosynthetic rate per mass & "photosynthetic rate per mass" OR "rate of CO2 assimilation per mass" OR "Amass" OR "Amax" \\
  Hmax & Maximum height & "potential height" OR "maximum height" OR "Hmax" \\
  Ks & Stem specific conductance & "Ks" OR "stem specific conductance" OR "stem hydraulic conductance" OR "sapwood conductance" \\
  LA & Leaf lamina area & "leaf length" OR "leaf lamina area" OR "leaf width" OR "leaf size" \\
  LAR & Leaf area ratio & "leaf area ratio" \\
  LMR & Leaf mass ratio & "leaf mass ratio" OR "leaf mass fraction" \\
  NARarea & Net assimilation rate per area &  "net assimilation rate per area" OR "rate of dry mass increament per area" OR "NARarea" OR "NAR" \\
  Narea & Leaf nitrogen content per area & "Narea" OR "nitrogen per area" OR "N/area" OR "leaf nitrogen content per area" OR "LNCarea" OR "leaf nitrogen content" \\
  NARmass & Net assimilation rate per mass &  "net assimilation rate per mass" OR "rate of dry mass increament per mass" OR "NARmass" OR "NAR" \\
  Nmass & Leaf nitrogen content per mass & "Nmass" OR "nitrogen per mass" OR "N/mass" OR "leaf nitrogen content per mass" OR "LNCmass" OR "leaf nitrogen content" \\
  Parea & Leaf phosphorus content per area & "Parea" OR "leaf P" OR  "leaf phosphorus content per area"  OR "leaf phosphorus content" \\
  Pmass & Leaf phosphorus content per mass & "Pmass" OR "leaf P" OR  "leaf phosphorus content per mass" OR "leaf phosphorus content" \\
  SA/LA & Sapwood area per leaf area & "sapwood area per leaf area" OR "huber value" OR "leaf area per sapwood area" \\
  Seed mass & Seed mass & "seed mass" OR "seed size" OR "seed volume" \\
  SLA & Specific Leaf Area & "leaf mass per area" OR "specific leaf area" OR "leaf construction cost" \\
  Thickness & Leaf thickness & "leaf thickness"  OR "leaf tissue density" \\
  Vessel density & Vessel density & "vessel density" OR "stem conduit density" \\
  Vessel size & Vessel area or diameter & "conduit area" OR "stem conduit area" OR "vessel diameter" OR "vessel size" OR "conduit size"   \\
  WD & Wood density & "wood density" OR"WD" OR "wood specific gravity" OR "stem density" OR "stem specific density" OR "SSD" \\
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics{output/FigA9.pdf}
\caption{\textbf{Funnel plots with filled in data based on the trim and fill method, showing transformed correlation coefficient z between growth and the traits a) Specific leaf area (SLA), b) Wood density (WD), c) Asymptotic height (Hmax), d) Seed mass and e) Assimilation rate of CO2 (Aarea)}. Shown here are ideal dataset (recorded across species and under unstressed conditions).  Note that in these plots the y axis is the 1/number of species included in the correlation, not the replication for the correlation. The number of species is sometimes as low as 1, and these are within-species correlations. The number of replicates used for the correlation coefficient was often much greater than the number of species. Nevertheless, the characteristic shape of funnel plots is observed, with correlations involving few species showing much wider variation even though within-species replication was sometimes high. The open points correspond to the estimated missing studies. The number of missing values is 9 for SLA,1 for WD, 0 for Hmax, 6 for Seedmass and 0 for Aarea}

@@

\begin{figure}[htbp]
\centering
\includegraphics{output/FigA10.pdf}
\caption{\textbf{Growth measurement used by traits and by size.} Shown here are complete dataset (recorded across  species and under both unstressed and stressed conditions). The colors correspond to the stage for which the r values have been recorded: red bars corresponds adult stage, orange bars to saplings, and grey bars to juveniles.}
\label{FigA10}
\end{figure}
